NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-31 paper/attention.tex
update 4 5

\subsection{Attention View}
\label{sec:attention}
As discussed in Section~\ref{sec:attention}, the attention is the only intermediate layer in the network that provides interpretable information for domain experts to infer the inner mechanisms of the model.
%
Intuitively, the attention captures the alignment of words between input sentences. For the NLI model examined in this work, the attention is represented as a matrix, in which the entries in the $i^{th}$ row correspond to the probabilities of words in the hypotheses aligned to the $i^{th}$ word in the premise.

>>>>

\subsection{Attention View}
\label{sec:attentionView}
As discussed in Section~\ref{sec:attention}, the attention is the only intermediate layer in the network that provides interpretable information for domain experts to infer the inner mechanisms of the model.
%
Intuitively, the attention captures the alignment of words between input sentences. For the NLI model examined in this work, the attention is represented as a matrix, in which the entries in the $i^{th}$ row correspond to the probabilities of words in the hypotheses aligned to the $i^{th}$ word in the premise.

### 2018-03-31 paper/overview.tex
space squeeze
update 4 5

\begin{figure}[htbp]
\centering
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{
 Perturbation-driven exploration of the natural language inference model.
 In the proposed tool, we enable the interrogation of the relationship between different components of the model via the perturbation-based analysis.
 %
 The user can perturb the input sentences (i.e., replace words with synonymous), perturb the attention (i.e., alter the soft alignment between sentences), and perturb the prediction (i.e., adjust the prediction by making updates to the model).
}
\label{fig:modelPipeline}
\end{figure}

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-3}).
%
As discussed in the previous section, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the all pair summary view (Section~\ref{sec:allPairs}).

>>>>

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-3}).
%
As discussed in the previous section, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (i.e., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the pair summary view (Section~\ref{sec:allPairs}).

\begin{figure}[htbp]
\vspace{-2mm}
\centering
 \includegraphics[width=1.0\linewidth]{pipeline}
 \vspace{-5mm}
 \caption{
 Perturbation-driven exploration of the natural language inference model.
 In the proposed tool, we enable the interrogation of the relationship between different components of the model via the perturbation-based analysis.
 %
 The user can perturb the input sentences (i.e., replace words with synonymous), perturb the attention (i.e., alter the soft alignment between sentences), and perturb the prediction (i.e., adjust the prediction by making updates to the model).
}
\vspace{-2mm}
\label{fig:modelPipeline}
\end{figure}

### 2018-03-31 paper/task.tex
update 4 5

%Through many rounds of discussions with NLP experts,
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
To support the visual exploratory workflow, we need to address the following visualization tasks.

@@

\item \textbf{T1}: Understand the stability of a prediction, i.e., how perturbation of input affects the behavior of the model?

\item \textbf{T2}: Examine the attention mechanism, i.e., what is the relationship between input and attention and how does the attention affect the prediction?

\item \textbf{T3}: When the predicted label is wrong, how can we update the model to correct the prediction? What is the effect of updating different parts of the model?

>>>>

%Through many rounds of discussions with NLP experts,
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
To support the exploratory workflow for analyzing NLI models, we need to address the following tasks.

@@

\item \textbf{T1}: Understand the stability of a prediction, i.e., how perturbations of the input affect the behavior of the model?

\item \textbf{T2}: Examine the attention mechanism, i.e., what is the relationship between the input and the attention, and how does the attention affect the prediction?

\item \textbf{T3}: When the predicted label is wrong, how can we update the model to correct the prediction? What are the effects of updating different parts of the model?
